### 2024-03-07 manuscript.tex
Reversed change in hierarchy.py file names for backwards compatibility
Fixed order of number for supporting figure

To better understand the relationship of this variability to differences in the regions' biophysical properties, we sought to correlate the average disorder score of the segments in a region with their molecular features. However, as the sequences are not independent but instead related by a hierarchical structure which reflects their evolutionary relationships, any features derived from them are unsuitable for direct use in many standard statistical procedures. In the most severe cases, traits derived from clades of closely-related species can effectively act as duplicate observations, which can yield spurious correlations. We therefore applied the method of contrasts to both the disorder scores and the features~\cite{Felsenstein1973, Felsenstein1985}. This algorithm takes differences between adjacent nodes in the phylogenetic tree relating the species to generate ``contrasts,'' which, under some general assumptions of the underlying evolutionary process, are independent and identically-distributed and therefore appropriate for use in correlation analyses. (A sample calculation using this algorithm is shown in Fig~\ref{sfig:contrasts}.) The resulting feature contrasts have varying degrees of correlation with the score contrasts (Fig~\ref{fig:scores}B-C). Some, like isopoint, are uncorrelated, but most are significantly, if weakly, correlated. In general, the strongest correlations are observed for features which have a direct biophysical relationship to the presence or absence of disorder, such as disorder\_fraction or hydrophobicity. Interestingly, the correlations with many motifs were statistically significant, though small in magnitude relative to the non-motif features. However, a more detailed analysis of this observation is presented in the discussion. To determine if regions with rapidly evolving disorder scores are associated with particular functions, processes, or compartments, we then extracted the regions in the upper decile of the rate distribution and performed a term enrichment analysis on their associated annotations (\ref{sfig:scores_histogram}~Fig). The most significant terms are generally related to DNA repair or extracellular structure, which suggests these processes and components are enriched in proteins whose structural state is rapidly evolving (Fig~\ref{fig:scores}E).

@@

\subsection{Clustering by evolutionary signatures reveals limited patterns of conserved features}
Given our observation of widespread signals of feature conservation in \textit{Drosophila} IDRs, we next sought to investigate if there are IDRs with similar patterns of conserved features and common functions. Accordingly, we treated the set of log likelihoods ratios for each region as its ``evolutionary signature'' and clustered them as in Zarin \textit{et al.}~\cite{Zarin2019}. The resulting heatmap reveals patterns of similar signatures interspersed among a high background of noise, and we identified at least 39 clusters with strong and consistent patterns of constraint (Fig~\ref{fig:hierarchy}). To gain greater insight into their functional properties, we performed term enrichment analyses using the annotations associated with each region's protein. After correcting for multiple comparisons using a 5\% false discovery rate, only five clusters were significantly enriched for any terms, which are given in Table~\ref{table:cluster_terms}. Most enriched terms are associated with fewer than ten proteins in each cluster, though the enriched terms for cluster 12 are markedly stronger and consistently associated with protein localization. We also observed that while many clusters are defined by a single strongly conserved feature, three (clusters 15, 23, and 24) have multiple features with signals of constraint. To determine if these more highly constrained regions are associated with a common underlying pattern of features, we compared their evolutionary signatures to normalized values of their inferred optima (Fig~\ref{fig:clusters}B-D). The heatmaps reveal that while the regions in these clusters have common patterns of constraint, the optimal values of their features can vary considerably. Likewise, cluster 12's strongly constrained feature, repeat\_K, is similarly variable (Fig~\ref{fig:clusters}A). Interestingly, this feature is consistent with its enrichment in annotations related to protein localization, as mitochondrial or nuclear targeting signals are characterized by a net positive charge or clusters of basic residues, respectively~\cite{Moulin2019, Lu2021}. However, the relative depletion of lysine repeats in some regions implies they are constrained against the formation of such targeting signals. This interpretation in turn suggests that IDRs may in general have complex relationships between their evolutionary constraints and functions where constraints can both preserve beneficial functions as well as prevent the acquisition of deleterious ones.

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{hierarchy/out/hierarchy_delta_loglikelihood.png}
\centering
\caption{\textbf{Hierarchical clustering of evolutionary signatures.}
Clusters are indicated by rectangles on the right. Clusters with fewer than 75 regions are omitted for clarity.}
\label{fig:hierarchy}
\end{figure}

>>>>

To better understand the relationship of this variability to differences in the regions' biophysical properties, we sought to correlate the average disorder score of the segments in a region with their molecular features. However, as the sequences are not independent but instead related by a hierarchical structure which reflects their evolutionary relationships, any features derived from them are unsuitable for direct use in many standard statistical procedures. In the most severe cases, traits derived from clades of closely-related species can effectively act as duplicate observations, which can yield spurious correlations. We therefore applied the method of contrasts to both the disorder scores and the features~\cite{Felsenstein1973, Felsenstein1985}. This algorithm takes differences between adjacent nodes in the phylogenetic tree relating the species to generate ``contrasts,'' which, under some general assumptions of the underlying evolutionary process, are independent and identically-distributed and therefore appropriate for use in correlation analyses. (A sample calculation using this algorithm is shown in~\ref{sfig:contrasts}~Fig.) The resulting feature contrasts have varying degrees of correlation with the score contrasts (Fig~\ref{fig:scores}B-C). Some, like isopoint, are uncorrelated, but most are significantly, if weakly, correlated. In general, the strongest correlations are observed for features which have a direct biophysical relationship to the presence or absence of disorder, such as disorder\_fraction or hydrophobicity. Interestingly, the correlations with many motifs were statistically significant, though small in magnitude relative to the non-motif features. However, a more detailed analysis of this observation is presented in the discussion. To determine if regions with rapidly evolving disorder scores are associated with particular functions, processes, or compartments, we then extracted the regions in the upper decile of the rate distribution and performed a term enrichment analysis on their associated annotations (\ref{sfig:scores_histogram}~Fig). The most significant terms are generally related to DNA repair or extracellular structure, which suggests these processes and components are enriched in proteins whose structural state is rapidly evolving (Fig~\ref{fig:scores}E).

@@

\subsection{Clustering by evolutionary signatures reveals limited patterns of conserved features}
Given our observation of widespread signals of feature conservation in \textit{Drosophila} IDRs, we next sought to investigate if there are IDRs with similar patterns of conserved features and common functions. Accordingly, we treated the set of log likelihoods ratios for each region as its ``evolutionary signature'' and clustered them as in Zarin \textit{et al.}~\cite{Zarin2019}. The resulting heatmap reveals patterns of similar signatures interspersed among a high background of noise, and we identified at least 39 clusters with strong and consistent patterns of constraint (Fig~\ref{fig:hierarchy}). To gain greater insight into their functional properties, we performed term enrichment analyses using the annotations associated with each region's protein. After correcting for multiple comparisons using a 5\% false discovery rate, only five clusters were significantly enriched for any terms, which are given in Table~\ref{table:cluster_terms}. Most enriched terms are associated with fewer than ten proteins in each cluster, though the enriched terms for cluster 12 are markedly stronger and consistently associated with protein localization. We also observed that while many clusters are defined by a single strongly conserved feature, three (clusters 15, 23, and 24) have multiple features with signals of constraint. To determine if these more highly constrained regions are associated with a common underlying pattern of features, we compared their evolutionary signatures to normalized values of their inferred optima (Fig~\ref{fig:clusters}B-D). The heatmaps reveal that while the regions in these clusters have common patterns of constraint, the optimal values of their features can vary considerably. Likewise, cluster 12's strongly constrained feature, repeat\_K, is similarly variable (Fig~\ref{fig:clusters}A). Interestingly, this feature is consistent with its enrichment in annotations related to protein localization, as mitochondrial or nuclear targeting signals are characterized by a net positive charge or clusters of basic residues, respectively~\cite{Moulin2019, Lu2021}. However, the relative depletion of lysine repeats in some regions implies they are constrained against the formation of such targeting signals. This interpretation in turn suggests that IDRs may in general have complex relationships between their evolutionary constraints and functions where constraints can both preserve beneficial functions as well as prevent the acquisition of deleterious ones. Furthermore, though a similar analysis of the inferred optima reveals some analogous patterns of clusters, they are generally more fragmented and diffuse, indicating that despite these caveats, the log likelihood ratios can highlight common patterns of constraint that are not readily apparent from the optimal values of the molecular features alone (\ref{sfig:hierarchy_mu}~Fig).

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{hierarchy/out/hierarchy.png}
\centering
\caption{\textbf{Hierarchical clustering of evolutionary signatures.}
Clusters are indicated by rectangles on the right. Clusters with fewer than 75 regions are omitted for clarity.}
\label{fig:hierarchy}
\end{figure}
